### 2021-06-29 outputs/paper/paper-copy.Rmd
Updates to scripts, fixes to 2020/2021 data, paper updates

---
title: "Reproducibility of COVID-19 pre-prints in 2020"
subtitle: "Markers of open code or data not found in 81 per cent of pre-prints"
author: 
  - Rohan Alexander^[University of Toronto.]
  - Annie Collins^[University of Toronto.]
thanks: "Collins thanks CANSSI Ontario for financial support. Code and data are available at: https://github.com/anniecollins/reproducibility."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "We create a dataset of all the pre-prints published on arXiv, bioRxiv, and medRxiv between 28 January 2020 and 31 January 2021. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. We are unable to find markers of either open data or open code for 81 per cent of the pre-prints on medRxiv  in our sample, and X per cent of those on arXiv in our sample. Our paper demonstrates the need to have authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, the need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE)
```

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
library(stargazer)
library(wesanderson)
library(kableExtra)
```

```{r loaddata, include=FALSE}
med_open_data_results <- 
  read_csv(here::here("outputs/data/med_open_data_results.csv"))
med_results <- 
  read_csv(here::here("outputs/data/med_results.csv")) # All COVID medRxiv papers (12,060)
med_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/med_2019_open_data_results.csv")) # All pre-prints posted to medRxiv in 2019

bio_open_data_results <- 
  read_csv(here::here("outputs/data/bio_open_data_results.csv"))
bio_results <- 
  read_csv(here::here("outputs/data/bio_results.csv")) # All COVID bioRxiv papers (3,714)
bio_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/bio_2019_open_data_results.csv"))
bio_2019_results <- 
  read_csv(here::here("outputs/control-data/bio_data_2019.csv"))

arxiv_open_data_results <- 
  read_csv(here::here("outputs/data/arxiv_open_data_results.csv"))
all_arxiv_results <- 
  read_csv(here::here("outputs/data/arxiv_results.csv")) # All COVID arXiv papers (3,812)

socarxiv_open_data_results <- 
  read_csv(here::here("outputs/data/socarxiv_open_data_results.csv"))
socarxiv_results <- 
  read_csv(here::here("outputs/data/socarxiv_results.csv")) # All COVID medRxiv papers (444). Same as socarxiv_sample.

all_arxiv_results <- 
  all_arxiv_results %>% 
  mutate(submitted = as_date(submitted),
         updated = as_date(updated))
```

# Introduction

Scientists use open repositories of papers to more quickly disseminate their research than is possible in traditional journals or conference proceedings. These repositories, such as arXiv, bioRxiv, medRxiv, and SocArXiv, are a critical component of scientific communication and many results build on the work published there. Thus it is important that the results that are published are credible. Pre-print repositories have drawn unprecedented attention in the context of the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community [@else2020]. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a pre-print

Neither peer-review nor credentials are a panacea nor a guarantee of quality. And the gate-keeping and slow publication times of traditional journals mean pre-print repositories have an important role to play. But given the importance of these repositories, it is important that scientists impose on themselves various standards for their results. Following @weissgerber2021automated we examine papers about COVID-19 published to  arXiv, bioRxiv, medRxiv, and SocArXiv from January 2020, through to May 1, 2021. We search for markers of open science as indicators of reproducibility, specifically open data and open code.

We find that approximately 76 per cent of papers sampled from arXiv, 68 percent of papers sampled from bioRxiv, 80 per cent of papers sampled from medRxiv, and 83 per cent of papers sampled from socArXiv contain neither open data nor open code markers. Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious increasing or decreasing trend throughout the pandemic. We also find that the presence of open data or open code markers has little influence on a pre-print's future publication, while the subset of sampled pre-prints that has been published contains an overall lower proportion of papers with these markers. Finally, we briefly examine the presence of open data and code markers in specific types of pre-prints, namely modelling, simulation, and machine learning research.

The remainder of this paper is structured as follows: in Section \@ref(methodology) we discuss the process of constructing our dataset through retrieving pre-prints from the arXiv, bioRxiv, medRxiv, and socArXiv repositories and mining them for open data and open code markers. In Section \@ref(results), we explore the results and key findings of this process. Finally, in Section \@ref(discussion) we discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

# Methodology

## Pre-print metadata
Our primary dataset consists of pre-print metadata extracted from the arXiv, bioRxiv, medRxiv, and socArXiv pre-print repositories via their respective Application Programming Interface (API). This metadata varies by repository, but generally includes the following information: title, abstract, author(s), date created, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, published DOI (if the pre-prints has since been published in a peer-reviewed journal), and download link. The data collection process was conducted separately for COVID-19 and pre-COVID-19 papers.

For COVID-19-related pre-prints, we first created a local copy of each repository containing all metadata for pre-prints registered between January 1, 2020 and May 1, 2021. In the case of socArXiv there are additionally fields that are self-reported by the submitting author about the availability of code and data. We classified individual pre-prints as "COVID-19-related" on the basis of containing one of more of the following terms in their title or abstract (case insensitive): "COVID-19", "COVID 19", "corona virus", "coronavirus", "coronavirus-2", "SARS-CoV-2", "SARSCoV-2", or "2019-nCoV". From each repository's dataset we randomly sampled pre-prints for further analysis, with the exception of socArXiv for which we analyzed all pre-prints that matched the above criteria due to their limited number (n = 410).

For pre-COVID-19 pre-prints, we created a local copy of each repository containing all metadata for pre-prints registered between January 1, 2019 and December 31, 2019. Since medRxiv was launched in June 2019, we used all pre-print data from the latter half of 2019. We then randomly sampled 1,200 pre-prints from each repository's dataset for analysis, with the exception of medRxiv for which only 913 pre-prints were available in total.

## Open data and code detection

We checked our sampled pre-prints for open data and code markers using the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub] within the `oddpub` R package [@citeODDPubpackage]. This required downloading each pre-print as a PDF and then converting the PDFs to text files. Some papers from socArXiv were only available in Word document format (.docx) and were downloaded as such and then converted to text format. We then conducted the open data and open code detection procedure which involved searching for keywords and other markers of open data and open code availability. A full listing of these markers is included in Appendix \@ref(markers). This was conducted using the `open_data_search()` function from the `oddpub` package.

The result of this process is a dataset indicating the presence of open data or open code markers in each pre-print (with a logical vector for each marker, followed by the relevant open data or open code statements where applicable). Our final dataset was formed by joining this output with the original sample metadata, typically using the DOI or the unique file name, to form a dataset including all original, qualitative metadata for each pre-print alongside its open data and open code status and markers.

Although socArXiv's metadata contained author-reported information concerning data and code availability for each pre-print, we elected to use the results of the ODDPub algorithm in our analysis to ensure comparability of results between pre-prints from all repositories. Notable discrepancies between the ODDPub algorithm and authors' self-reported data and code availability include...

```{r monthly-papers-total, fig.cap = "Number of pre-prints related to COVID-19", out.width="90%", fig.align = "center"}
just_arxiv <- 
  all_arxiv_results  %>% 
  select(submitted) %>% 
  mutate(type = "arXiv") %>% 
  rename(date = submitted)

just_medRxiv <- 
  med_results  %>% 
  select(date) %>% 
  mutate(type = "medRxiv")

just_bioRxiv <- 
  bio_results  %>% 
  select(date) %>% 
  mutate(type = "bioRxiv")

just_socarxiv <- 
  socarxiv_results  %>% 
  select(date_created) %>%
  rename(date = "date_created") %>%
  mutate(type = "socArXiv")

all <- rbind(just_arxiv, just_medRxiv, just_bioRxiv, just_socarxiv)

all %>% 
  mutate(month = floor_date(date, "month")) %>% 
  filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month)) + 
  geom_histogram(aes(fill = type), stat="count", position = "dodge") + 
  labs(title = "COVID-19-related pre-prints posted per month",
       subtitle = "Total of 3,896 relevant pre-prints on arXiv, 3,714 on bioRxiv, 12,060 on medRxiv, 410 on socArXiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1"))
```

# Results

## All pre-prints
The number of pre-prints posted per month increased dramatically in the first half of 2020 across all repositories, reaching its maximum sometime between April and June (depending on repository) and subsequently decreasing. The number of pre-prints posted monthly since August 2020 has remained reasonably steady, with the exception of medRxiv which experienced a spike up to nearly 1000 pre-prints posted in March 2021 (Figure \@ref(fig:monthly-papers-total)). For context, COVID-19 was declared a pandemic by the World Health Organization (WHO) on March 11, 2020, at which point the number of cases globally had just surpassed 118,000 (primarily in east Asia) and the virus had been reported in 114 countries [@citeWHOtimeline].

## Open data and code
We analyzed a total of 3,910 COVID-19-related pre-prints. This sample is further broken down as follows: 1,500 from medRxiv, 1,000 from arXiv, 1,000 from bioRxiv, and 410 from socArXiv. Broadly, we are unable to find markers of either open data for open code for 2,987 or approximately 76 per cent of our sample (Table \@ref(tab:summarycounts-overall)). Of the remaining pre-prints, 6 per cent contained open code markers only, 10 per cent contained open data markers only, and 8 percent included markers of both open data and open code.

Once differentiated by repository, we observe that open data and code markers were absent from 68 percent of the sampled bioRxiv pre-prints, 80 per cent of the sampled medRxiv pre-prints, 76 per cent of the sampled arXiv pre-prints, and 83 per cent of all COVID-19-related socArXiv pre-prints. The distribution of the remaining portion of pre-prints varies by repository (Table \@ref(tab:summarycounts-repositories)). Notably, 28 per cent of sampled pre-prints from bioRxiv contained open data markers and 20 per cent of sampled arXiv pre-prints contained markers of open code, the highest proportions of any repository for each type of marker.

```{r, summarycounts-overall}
just_arxiv <-
  arxiv_open_data_results  %>%
  select(submitted, is_open_data, is_open_code) %>%
  mutate(submitted = as.Date(submitted)) %>%
  mutate(type = "arXiv") %>%
  rename(date = submitted)

just_medRxiv <-
  med_open_data_results  %>%
  mutate(date = as.Date(date))  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(type = "medRxiv")

just_bioRxiv <-
  bio_open_data_results  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(date = as.Date(date)) %>%
  mutate(type = "bioRxiv")

just_socarxiv <-
  socarxiv_open_data_results  %>%
  select(date_created, is_open_data, is_open_code) %>%
  rename(date = "date_created") %>%
  mutate(type = "socArXiv")

all <- rbind(just_arxiv, just_medRxiv, just_bioRxiv, just_socarxiv)

all %>%
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>%
  select(is_open_data, is_open_code, n, prop_total) %>%
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Counts and proportions of open data and code markers in sample",
        booktabs = TRUE,
        digits = 2)
```

```{r summarycounts-repositories}
# open data/open code count summary table

all %>%
  group_by(type) %>%
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>%
  ungroup() %>%
  select(is_open_data, is_open_code, n, prop_total) %>%
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Counts and proportions of open data and code markers by pre-print repository",
        booktabs = TRUE,
        digits = 2) %>%
  group_rows(index = c("arXiv" = 4, "bioRxiv" = 4, "medRxiv" = 4, "socArXiv" = 4))
```

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \@ref(fig:monthly-papers-condition-stack)). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to any given repository (Figure \@ref(fig:monthly-proportion-of-open)). In our datasets, very few (if any) pre-prints were sampled for the month of January 2020. None of these pre-prints contained open data or open code markers, thus the 0 per cent rate of open data and code for this month across all repositories should be considered an outlier.

```{r monthly-papers-condition-stack, fig.cap = "Number of pre-prints related to COVID-19 sampled from medRxiv, distinguished by presence of open data or code markers", out.width="90%", fig.align = "center"}
# Number of papers for each open data/code status per month
open_status_med <- 
  med_open_data_results %>% 
  select(title, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "medRxiv")

open_status_bio <- 
  bio_open_data_results %>% 
  select(title, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "bioRxiv", date = as.character(date))

open_status_arxiv <- 
  arxiv_open_data_results %>% 
  select(title, submitted, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "arXiv", date = as.character(submitted) %>% str_sub(end = 10)) %>%
  select(title, date, is_open_code, is_open_data, condition, repo)

open_status_socarxiv <- 
  socarxiv_open_data_results %>% 
  select(title, date_created, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "socArXiv", date = as.character(date_created)) %>%
  select(title, date, is_open_code, is_open_data, condition, repo)

open_status <- rbind(open_status_med, open_status_bio, open_status_arxiv, open_status_socarxiv)

open_status$date <- as.Date(open_status$date)
open_status$condition[open_status$condition == 2] <- "Both"
open_status$condition[open_status$condition == 0] <- "Neither"
open_status$condition[open_status$condition == 1 & open_status$is_open_code == 1] <- "Open Code"
open_status$condition[open_status$condition == 1 & open_status$is_open_data == 1] <- "Open Data"

open_status %>% 
  ggplot() + 
  geom_bar(aes(x = factor(floor_date(date, "month")), 
              fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")), 
              position = "stack")) +
  labs(fill = "Type of marker") +
  scale_x_discrete(breaks=c("2020-01-01","2020-07-01", "2021-01-01"), 
                   labels=c("Jan 2020", "Jul 2020", "Jan 2021")) +
  ylab("Number of pre-prints") +
  xlab("Month") +
  labs(title="Sampled pre-prints by month", subtitle="medRxiv, Total = 1,200") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  facet_wrap(~repo)
```

```{r monthly-proportion-of-open, fig.cap= "Proportion of pre-prints in sample with open data or open code markers by month", out.width="90%", fig.align = "center"}
# Proportion of papers posted with open data or open code each month
med_month_sum_cov <- 
  med_open_data_results %>% 
  mutate(date = as.Date(date)) %>%
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count = n(), "Open Code" = mean(is_open_code), "Open Data" = mean(is_open_data)) %>% 
  mutate(daily_rate = count/days_in_month(month), repo = "medRxiv")

bio_month_sum_cov <- 
  bio_open_data_results %>% 
  mutate(date = as.Date(date)) %>%
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count = n(), "Open Code" = mean(is_open_code), "Open Data" = mean(is_open_data)) %>% 
  mutate(daily_rate = count/days_in_month(month), repo = "bioRxiv")

arxiv_month_sum_cov <- 
  arxiv_open_data_results %>% 
  group_by(month = floor_date(submitted, "month")) %>% 
  summarise(count = n(), "Open Code" = mean(is_open_code), "Open Data" = mean(is_open_data)) %>% 
  mutate(daily_rate = count/days_in_month(month), repo = "arXiv")

socarxiv_month_sum_cov <- 
  socarxiv_open_data_results %>% 
  group_by(month = floor_date(date_created, "month")) %>% 
  summarise(count = n(), "Open Code" = mean(is_open_code), "Open Data" = mean(is_open_data)) %>% 
  mutate(daily_rate = count/days_in_month(month), repo = "socArXiv")

month_sum_cov <- rbind(med_month_sum_cov, bio_month_sum_cov, arxiv_month_sum_cov, socarxiv_month_sum_cov) %>%
  pivot_longer(cols = c("Open Data", "Open Code")) %>%
  filter(value != 1) # Removes outliers where small number of papers in a month was sampled and both had open data/code

# TODO: Fix date labels, facet_wrap labels
month_sum_cov %>%
  ggplot() +
  geom_line(data = month_sum_cov, aes(month, value, color = repo)) +
  ggtitle("Proportion of sampled pre-prints with open data or code markers") +
  labs(colour = "Repository",
       x = "Month",
       y = "Proportion of pre-prints") + 
  theme_minimal() +
  scale_color_manual(values = wes_palette("Darjeeling1")) +
  facet_wrap(~name)

```

## Publication status

Our metadata also contains an indication of whether or not a pre-print has been published in the form of a DOI linking to the corresponding published paper. The proportion of pre-prints that have been published varies by repository (Table \@ref(tab:published-sample-summary)). Notably, of all COVID-19-related pre-prints created during 2020 and 2021, approximately one third of those posted to bioRxiv and nearly 30 percent of those posted to medRxiv were eventually published in peer reviewed journals. This is high in comparison to the proportions from arXiv and socArXiv, which may be a reflection of the unprecedented attention pre-print servers showcasing research from the biomedical and life sciences have received throughout the pandemic.

```{r published-sample-summary}
med_published <- 
  med_results %>% 
  select(title, date, published) %>% 
  mutate(repo = "medRxiv", published = as.numeric(!is.na(published)))

bio_published <- 
  bio_results %>% 
  select(title, date, published) %>% 
  mutate(repo = "bioRxiv", published = as.numeric(!is.na(published)))

arxiv_published <- 
  all_arxiv_results %>% 
  select(title, submitted, doi) %>%
  mutate(doi = as.numeric(!is.na(doi)), repo = "arXiv") %>%
  rename(published = doi, date = submitted)

socarxiv_published <- 
  socarxiv_results %>% 
  select(title, date_created, links.doi) %>%
  mutate(links.doi = as.numeric(!is.na(links.doi)), repo = "socArXiv") %>%
  rename(published = links.doi, date = date_created)

published <- rbind(med_published, bio_published, arxiv_published, socarxiv_published)

published %>% 
  group_by(repo) %>% 
  summarise(`Published pre-prints` = sum(published), `Proportion published` = mean(published) %>% round(2)) %>%
  kable(caption = "Counts and proportion of published COVID-19 pre-prints in each repository",
        col.names = c("Repositiory", "Published pre-prints", "Proportion published"))
```

In Table \@ref(tab:open-published-summary) we disaggregate sampled pre-prints by whether or not there is an indication of publication. We find that the proportion of pre-prints with open data or code markers among those that have been published is approximately the same as pre-prints that have not been published, differing by one percent only within the bioRxiv and arXiv samples, and three percent in the socArxiv sample.

```{r open-published-summary}
# Maybe find a way to use the data frames constructed above instead of constructing new ones?

# ArXiv variables: doi
# SocArXiv variables: links.doi

med_published_open_status <- 
  med_open_data_results %>% 
  select(title, date, published, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "medRxiv")

bio_published_open_status <- 
  bio_open_data_results %>% 
  select(title, date, published, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "bioRxiv")

arxiv_published_open_status <- 
  arxiv_open_data_results %>% 
  select(title, submitted, doi, is_open_code, is_open_data) %>%
  mutate(doi = as.numeric(!is.na(doi)), condition = is_open_data + is_open_code, repo = "arXiv") %>%
  rename(published = doi, date = submitted)

socarxiv_published_open_status <- 
  socarxiv_open_data_results %>% 
  select(title, date_created, links.doi, is_open_code, is_open_data) %>%
  mutate(links.doi = as.numeric(!is.na(links.doi)), condition = is_open_data + is_open_code, repo = "socArXiv") %>%
  rename(published = links.doi, date = date_created)

published_open_status <- rbind(med_published_open_status, bio_published_open_status, arxiv_published_open_status, socarxiv_published_open_status)

published_open_status$condition[published_open_status$condition == 2] <- "Both"
published_open_status$condition[published_open_status$condition == 0] <- "Neither"
published_open_status$condition[published_open_status$condition == 1 & published_open_status$is_open_code == 1] <- "Open code"
published_open_status$condition[published_open_status$condition == 1 & published_open_status$is_open_data == 1] <- "Open data"

published_open_status %>%
  count(repo, published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  mutate(Published = ifelse(published == 0, "No", "Yes"), `Proportion with neither`= round((`Neither`)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  select(Published, Both, Neither, `Open code`, `Open data`, `Proportion with neither`) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published",
        booktabs = TRUE) %>%
  group_rows(index = c("bioRxiv" = 2, "medRxiv" = 2, "arXiv" = 2, "socArXiv" = 2))
```

```{r open-published-stack, fig.cap= "Number of pre-prints in sample with open data or open code markers by publication status", out.width="90%", fig.align = "center"}

# TODO: are these graphs even necessary?

# published_open_status$published[published_open_status$published == 0] <- "Published"
# published_open_status$published[published_open_status$published == 1] <- "Unpublished"

published_open_status %>% 
  mutate(published = ifelse(published == 0, "Unpublished", "Published")) %>%
  ggplot() + 
  geom_bar(aes(x = published, fill=factor(condition, levels = c("Neither", "Open code", "Open data", "Both")), position = "stack")) +
  labs(fill = "Type of marker",
       y = "Number of pre-prints",
       x = "Publication status") +
  ggtitle("Pre-prints by publication status") +
  theme_minimal() +
  facet_wrap(~repo) + 
  scale_fill_manual(values = wes_palette("Darjeeling1"))

```

It is important to note that our dataset imperfectly characterizes publication. In particular, it likely does not have the publication details for some papers that were published. And even if it were a perfect record, there is a publication lag (estimated at an average of around 60 days for COVID-19-related pre-prints) that may skew the results for the papers in the latter portion of our sample [@kwon2020].

## Pre-pandemic Pre-prints

In order to examine the influence of the COVID-19 pandemic on open science practices in during the pandemic, we analyzed pre-prints posted between January and December 2019 from each of the four repositories using the same process as was done for the COVID-19-related pre-prints. Since medRxiv was founded in June 2019, all pre-prints unrelated to COVID-19 posted from the latter half of 2019 were analyzed (a total of 913). For bioRxiv, a random sample of 1,200 was taken from all non-COVID-19-related pre-prints posted in the relevant date range.

```{r 2019-results}
bio_2019 <- bio_2019_results %>%
  select(date) %>%
  mutate(type = "bioRxiv")

med_2019 <- med_2019_open_data_results %>%
  select(date) %>%
  mutate(type = "medRxiv")

all_2019 <- rbind(bio_2019, med_2019)

# TODO: position = "dodge" not working
all_2019 %>% 
  mutate(month = floor_date(date, "month")) %>% 
  # filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month, fill = type)) + 
  geom_histogram(stat="count", position = "dodge2") +
  labs(title = "Pre-prints posted per month, 2019",
       subtitle = "Total of 31,752 pre-prints on bioRxiv, 913 on medRxiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1"))
```

```{r 2019-open-results}
just_2019_medRxiv <-
  med_2019_open_data_results  %>%
  mutate(date = as.Date(date))  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(type = "medRxiv")

just_2019_bioRxiv <-
  bio_2019_open_data_results  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(date = as.Date(date)) %>%
  mutate(type = "bioRxiv")

all_2019 <- rbind(just_2019_medRxiv, just_2019_bioRxiv)

all_2019 %>%
  group_by(type) %>%
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>%
  ungroup() %>%
  select(is_open_data, is_open_code, n, prop_total) %>%
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Counts and proportions of open data and code markers in 2019 medRxiv and bioRxiv samples",
        booktabs = TRUE,
        digits = 2) %>%
  group_rows(index = c("bioRxiv" = 4, "medRxiv" = 4))
```

Between June and December 2019, the number of pre-prints posted to medRxiv monthly saw an overall increase which may be expected as the repository gained recognition and popularity in the medical research community. The number of papers posted monthly to bioRxiv also saw an overall increase throughout 2019. It is important to note here that due to its relative immaturity at the beginning of the COVID-19 pandemic, a significant portion of medRxiv's overall usage has been dedicated to COVID-19-related research. In total, 19,638 pre-prints have been posted to medRxiv since June 2019, 12,060 (approximately 61 per cent) of which relate to COVID-19.

Of the analyzed pre-prints from 2019, 63 percent of those posted to bioRxiv and 75 per cent of those posted to medRxiv showed no indication of open data or open code (Table \@ref(tab:2019-published)). Both of these rates are approximately five per cent lower than the corresponding proportions of COVID-19-related pre-prints, suggesting that the analyzed pre-prints from 2019 contained an overall higher prevalence of open data and code markers than pre-prints concerning COVID-19. Specifically, we found that open data availability in medRxiv pre-prints was significantly associated with a pre-pandemic registration date (chi-squared = 8.225, p < 0.005), as was open code availability for bioRxiv pre-prints (chi-squared = 15.028, p < 0.005). This would suggest that open data and code practices have suffered in the context of COVID-19.

```{r 2019-published}
med_published_2019 <- 
  med_2019_open_data_results %>% 
  select(title, date, published) %>% 
  mutate(repo = "medRxiv", published = as.numeric(published))

bio_published_2019 <- 
  bio_2019_results %>% 
  select(title, date, published) %>% 
  mutate(repo = "bioRxiv", published = as.numeric(!is.na(published)))

published_2019 <- rbind(med_published_2019, bio_published_2019)

published_2019 %>% 
  group_by(repo) %>% 
  summarise(`Published pre-prints` = sum(published), `Proportion published` = mean(published) %>% round(2)) %>%
  kable(caption = "Counts and proportion of published COVID-19 pre-prints in each repository, 2019",
        col.names = c("Repositiory", "Published pre-prints", "Proportion published"))
```

```{r, 2019-open-publishing}
med_2019_published_open_status <- 
  med_2019_open_data_results %>% 
  select(title, date, published, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "medRxiv")

bio_2019_published_open_status <- 
  bio_2019_open_data_results %>% 
  select(title, date, published, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "bioRxiv")

published_2019_open_status <- rbind(med_2019_published_open_status, bio_2019_published_open_status)

published_2019_open_status$condition[published_2019_open_status$condition == 2] <- "Both"
published_2019_open_status$condition[published_2019_open_status$condition == 0] <- "Neither"
published_2019_open_status$condition[published_2019_open_status$condition == 1 & published_2019_open_status$is_open_code == 1] <- "Open code"
published_2019_open_status$condition[published_2019_open_status$condition == 1 & published_2019_open_status$is_open_data == 1] <- "Open data"

published_2019_open_status %>%
  count(repo, published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  mutate(Published = ifelse(published == 0, "No", "Yes"), `Proportion with neither`= round((`Neither`)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  select(Published, Both, Neither, `Open code`, `Open data`, `Proportion with neither`) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published, 2019",
        booktabs = TRUE) %>%
  group_rows(index = c("bioRxiv" = 2, "medRxiv" = 2))
```

Examining publication rates for pre-pandemic papers, we observe that 64 per cent of pre-prints posted to bioRxiv and 61 per cent of pre-prints posted to medRxiv during 2019 were eventually peer reviewed and published (Table \@ref(tab:2019-published)). These rates are approximately double those of the COVID-19-related pre-prints posted to servers during 2020 and 2021 (33 per cent and 31 per cent from bioRxiv and medRxiv respectively). When disaggregated by open data and code status, we find that published and unpublished pre-prints contain open data and code markers in similar proportions (Table \@ref(tab:2019-open-publishing)). This mimics the behaviour of published and unpublished COVID-19-related pre-prints.

# Discussion

- Add reference somewhere to that paper about the numbers: https://link.springer.com/article/10.1007/s11192-020-03675-3
- Add reference somewhere to that paper about keyword search and then analysis. Was it a Gary King one?

## On the role of transparency and reproducibility

Transparency and reproducibility are hallmarks of quality scientific research. Open data and open code contribute greatly to both in allowing the scientific community to more easily verify the authenticity of purported scientific discovery and its supporting evidence, which is especially important in cases where scientific research may quickly and directly impact clinical practice or policy such as during the COVID-19 pandemic. Among a myriad of other impacts on biomedical research, COVID-19 has dramatically increased the popularity of pre-prints from both a production and consumption stand-point. The number of COVID-19 pre-prints posted to medRxiv increased dramatically in the early stages of the pandemic while non-COVID-19 pre-print rates remained consistent. The same trends were apparent in abstracts accessed by medRxiv users, where COVID-19 pre-print abstracts were viewed over 15 times the rate of non-COVID-19 pre-print abstracts [@citeFraser]. For these reasons, it is more important now than ever to examine open science standards and reproducibility within pre-print repositories.

Open data is generally accepted to be beneficial to the scientific process, and to a paper's reproducibility potential, hence it is concerning that over 75 per cent of pre-prints in our sample contained no open data markers. This concern is slightly mitigated by recognition of challenges in working with biomedical data compared with data in other fields, notably privacy and ethics concerns when working with personal data [@Floca2014]. The COVID-19 pandemic has seen an uptake in open science practices globally, as evidenced by the creation of open data repositories such as the dashboard maintained by the Center for Systems Science and Engineering at Johns Hopkins University [@citeCSSE] or the large number of publishers who have removed paywalls from published COVID-19 research [@gill2020]. (MENTION/REFERENCE https://wellcome.org/coronavirus-covid-19/open-data ?) 
- Insert discussion on difference from 2019 data

Open code as an open science marker is much more context and field-dependent, as not all biomedical research papers will rely on computational methods for their analyses. However, in pre-prints where code comprises a large portion of the methodology or results, posting it openly to repositories like GitHub contributes greatly to a pre-print's potential reproducibility. This gains importance as computational methods become increasingly popular in the rush to form predictions about emerging situations with limited data or laboratory research, which was the case for modelling studies in the early days of the COVID-19 pandemic (CITE). We also see growing concern over the quality and consequences of this sort of research, with bioRxiv no longer allowing purely computational work [@kwon2020].

## The role of of pre-print repositories

Many concerns have arisen from the unprecedented rate at which COVID-19 research has been posted and consumed via pre-print repositories, particularly in the early stages of the pandemic [@raynaud2021]. Any rushed scientific research has potential to skip (or at least place less precedence on) quality open science practices, thus it may be reasonable to expect a decrease of open data or code markers in the pre-prints posted during times of increased overall posting to medRxiv. In our analysis, we found little relationship between time posted and likelihood of having open data or code markers with the proportion of papers containing these markers fluctuating greatly from month to month and no apparent decrease during periods of increased publication. This suggests that open science practices are more highly influenced by other factors, perhaps training, publication bias, or the nature of the paper itself. On the other hand, we do not see an overall long-term increase in either open data or open code markers throughout our year of analysis which we may expect in the context of the aforementioned open science movements the pandemic has fostered. Although not pre-print specific, @else2020 found that overall research output has fluctuated between different fields and topics (namely modelling disease spread, public health, diagnostics and testing, mental health, and hospital mortality) throughout different stages of the pandemic which may account for some of the fluctuation and overall lack of noticeable trend over the course of the year.

To emphasize the ongoing need for open data and code in modelling a pandemic, we consider two high profile epidemiological models that emerged in early 2020. Modelling was conducted by Imperial College London (ICL) [@ferguson2020] and the Institute for Health Metrics and Evaluation (IHME) at the University of Washington [@murray2020], and both papers were initially posted to pre-print repositories. The ICL model went on to become the most cited pre-print as of December 2020 [@else2020], and both had significant influence over policy and public health decisions worldwide [@adam2020]. An independent review of these two models by @jin2020 found that while code and data were openly available for both, only the ICL model was reproducible due to limited transparency on the underlying methodology of the IHME model. The open-source nature of these papers was fundamental to reproduction attempts and is a great example of the need for open data and code in reproducing COVID-19 research, particularly in evaluating pre-prints as they begin to influence public decision-making.

In the context of the above factors, it was disheartening in our analysis to find that the proportion of modelling- and simulation-related pre-prints with open data or code increased only marginally from that of the entire sample. One might hope that modelling papers should universally be subject to the same analysis as conducted by @jin2020 as for the examples above, which is made possible by the availability of relevant code and data. Although our analysis was not particularly robust, this shows a need for future investigation and potential overall improvement in open science standards for these types of pre-prints (of course subject to the data and code considerations already discussed). This need is again emphasized by the new-found speed at which pre-print articles may gain public, media, and political attention in the context of the pandemic.

## The importance of open data and open code

Beyond pre-prints, COVID-19 has had great influence over publication and peer review processes as well, with expedited review timelines for COVID-19 papers at the expense of longer waits for other scientific research [@else2020]. Needless to say, it is important that open data and code standards be maintained in published work as well. Our findings in this regard were two-fold: that open data or code markers do not appear to directly influence a pre-print's eventual publication, and that a lower proportion of published papers contain either open data or code markers than the general sample (TODO: no longer valid; cite new McGuinness/SHephard paper). Both of these raise concerns over publication bias, the potential that journals have favored novel yet less transparent or reproducible papers over those with null results but a high standard of open science practices. Concerns have already been raised through systemic reviews of COVID-19 publications [@raynaud2021], and oversights in data accessibility have led to high profile retractions of publications in the past, for example two papers from The Lancet and the New England Journal of Medicine that were withdrawn due to concerns over the private nature of their underlying dataset [@ledford2020].

In all fields of science, increasing access to data and code used for pre-printed or published research is a step in the direction of more transparent, reproducible, and reliable research. The ongoing COVID-19 pandemic has created a novel, constantly changing scientific culture that should be navigated with the utmost care so as to uphold standards of scientific practice for both the research community and the safety of the general public. Our analysis shows that there is much improvement to be made in the areas of open data and code availability within COVID-19 pre-print papers on arXiv, bioRxiv, medRxiv, and socArXiv.

Demand for timely research / results in high frequency as epidemic is rapidly evolving. Preprints are more efficient at doing this because there is no time spent on peer review. However, they also allow lesser-known researchers to better disperse their research, because of the possibility that fast-tracked peer review may be biased towards established researchers. While there is a clear need for pre-prints, the point remains that they do not go through the peer review process. This question of quality and validity is particularly pertinent in the COVID-19 context because poorly validated results and false information may spread quickly, and have real effects. We are not saying that peer review implies that a paper is of a high-quality, we are instead saying that the provision of code and data alongside the preprint goes some way to allowing others to trust the findings of pre-prints even though they have not been peer-reviewed. One way this could be encouraged would be for pre-print repositories to have authors characterize the extent to which they have adopted open science practices as part of their submission. Although those papers that do not adopt these practices should not be rejected from pre-print repositories, greater clarity around this would be useful and might move the state-of-the-art forward.

## Weaknesses and next steps

In this paper we consider only pre-prints from medRxiv, but this analysis can and should be extended to other pre-print repositories including bioRxiv and arXiv. The `medrxivr` package [@citemedrxivr] can be used with the bioRxiv API as well, and our code can be modified to conduct the same scraping and text mining procedure as was used for medRxiv data.

We also wish to expand our analysis to consider the geographic distribution of research and the potential influence of different practices and policies concerning open science as pre-prints vary by location. This is pertinent to our current paper as the epicenter of the virus spread (and thus of scientific output) has shifted throughout the pandemic which has implications for our time-based analysis.

An important weakness to note is the potential presence of false negatives in indicators of publication in our dataset. @abdill2019 estimate that the false-negative rate may be as high as 37.5 per cent for data pulled from the bioRxiv API, meaning analysis of published papers may represent only a fraction of those that have actually been published. It is unclear to what extent this is the case for medRxiv or what bias may exist in the subset of pre-prints for which publication was detected, as it is likely that this process relies on title-based text matching [@abdill2019]. It is also likely that some of our more recent sampled pre-prints will be published in future which we could not account for at the time of our data collection.

We also recognize that this analysis relies heavily on text-based analysis which was not verified directly in most cases and may lead to higher levels of uncertainty. In future, we wish to take smaller sub-samples to validate factors like publication status or paper topic beyond simple keyword searches or API output.

\newpage

# (APPENDIX) Appendix {-}

# Markers

Insert details on what markers the `oddpub::open_data_search()` uses. Include info here directly from package documentation (i.e. keywords/phrases used for text parsing)?

# Classification

Insert details on keyword search used for machine learning/modelling/simulation

\newpage

# References